Analysis methods for characterizing the climate sensitivity of annual growth and woody productivity using tree-rings

### 2020-05-15 doc/methods/analysis_methods.Rmd
Update analysis_methods.Rmd

1.1. "Detrend" individual tree growth records (radial growth increments or $\Delta AGB$) by fitting a spline using a GAM, thereby producing residuals. This parallels detrending of traditional dendro analyses, which seems necessary to pull out a good climate signal (see "Considerations and alternative approaches"). However, rather than averaging by species as per traditional dendro methods, we will analyze individual-level residuals.

1.2. We use climate data from CRU v. 4.04. Based on how variables are related to one another, and how they are derived in CRU [Harris et al. 2020](https://rdcu.be/b3U3e), we grouped variables as follows, using climwin to eselct the best variable and time window for each: 
- Precip group (describes precip and its frequency): PRE, WET
- temperature group (describes temperature): TMP, TMIN, TMAX, PET
- temperature range group (describes variation in temperature, linked to cloud cover): DTR, CLD, PET

1.3. Use the [climwin](https://cran.r-project.org/web/packages/climwin/index.html) R package [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590), to identify the climate variables and time windows that most strongly influences tree growth *for all species combined*. 
Climwin can check for linear and quadratic relationships. We checked quadratic only because (i) it speeds up the process, (ii) preliminarly tests revealled that quadratic was usually the one with lower AIC, and when not there tended to be little difference in AIC and the curve would look like a straight line anyway, (iii) quadratic makes more biological sense, (iv) it would be too tricky to code the gam to adapt to any circumstance.
We put the residuals derived in step 1.1 in *climwin* with a mixed effects model using species and tree identity as random effect: `ln[residual] ~ [climate] +  + (1 | sp) + (1 | treeID)`. 
Outputs are saved [here](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/blob/master/results/ALL_species_mixed_model_on_residuals/). Example output is shown in Fig. 1.

@@

(*Note: This process will tend to pick out current growing season variables, but it may be that once we have 1-2 variables for current growing season, others would be more important (e.g., previous growing season, start/ end of growing season). Thus, we may want to modify how we select climate variables. ([GitHub issue # 19])(https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/19)*)

>>>>

1.1. "Detrend" individual tree growth records ($log[radial.growth.increment]$ or $log[\Delta AGB]$) by fitting a spline using a GAM, thereby producing residuals. This parallels detrending of traditional dendro analyses, which seems necessary to pull out a good climate signal (see "Considerations and alternative approaches"). However, rather than averaging by species as per traditional dendro methods, we will analyze individual-level residuals.

1.2. We use climate data from CRU v. 4.04. Based on how variables are related to one another, and how they are derived in CRU ([Harris et al. 2020](https://rdcu.be/b3U3e)), we grouped variables as follows, using climwin to select the best variable and time window for each: 
- Precip group (describes precip and its frequency): PRE, WET
- temperature group (describes temperature): TMP, TMIN, TMAX, PET
- temperature range group (describes variation in temperature, linked to cloud cover): DTR, CLD, PET

1.3. Use the [climwin](https://cran.r-project.org/web/packages/climwin/index.html) R package [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590), to identify the climate variables and time windows that most strongly influences tree growth *for all species combined*. 
Climwin can check for linear and quadratic relationships. We checked quadratic only because (i) it speeds up the process, (ii) preliminarly tests revealled that quadratic was usually the one with lower AIC, and when not there tended to be little difference in AIC and the curve would look like a straight line anyway, (iii) quadratic makes more biological sense, (iv) it would be too tricky to code the gam to adapt to any circumstance.
We put the residuals derived in step 1.1 in *climwin* with a mixed effects model using species and tree identity as random effect: `residual ~ [climate] + (1 | sp) + (1 | treeID)`. 
Outputs are saved [here](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/blob/master/results/ALL_species_mixed_model_on_residuals/). Example output is shown in Fig. 1.

@@

(*Note: This process will tend to pick out current growing season variables, but it may be that once we have 1-2 variables for current growing season, others would be more important (e.g., previous growing season, start/ end of growing season). Thus, we may want to modify how we select climate variables. ([GitHub issue # 19](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/19))*)
